A Detailed Analysis of Performance Bottlenecks in MRI Pre-Processing

### 2023-11-30 paper/paper.tex

\section{Methods}
\subsection{Infrastructure}
For our experiments, we used the \textit{Slashbin cluster} at Concordia University. The cluster has eight compute nodes, each configured with two 16-core Intel(R) Xeon(R) Gold 6130 CPU @ \SI{2.10}{\giga\hertz}, \SI{250}{\gibi\byte} of RAM, \SI{126}{\gibi\byte} of tmpfs, six \SI{447}{\gibi\byte} SSDs with the XFS file system (no RAID enabled), CentOS~8.1 and Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64}. \TG{I would defer this subsection to later in the methods as it's not the most central one}

@@

\begin{figure*}[ht!]
	\centering
	% ANTS brainExtraction
	\begin{subfigure}[t]{0.49\textwidth}
		\caption{}
		\label{subfig:hotspots-ants-brainExtraction}
		\includegraphics[width=\textwidth]{figures/hotspots-1thread-ants-brainExtraction.png}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.49\textwidth}
		\caption{}
		\label{subfig:hotspots-ants-brainExtraction-fp}
		\includegraphics[width=\textwidth]{figures/hotspots-1thread-ants-brainExtraction-fp.png}
	\end{subfigure}

	% ANTS registrationSyN
	\begin{subfigure}[t]{0.49\textwidth}
		\caption{}
		\label{subfig:hotspots-ants-registrationSyN}
		\includegraphics[width=\textwidth]{figures/hotspots-1thread-ants-registrationSyN.png}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.49\textwidth}
		\caption{}
		\label{subfig:hotspots-ants-registrationSyN-fp}
		\includegraphics[width=\textwidth]{figures/hotspots-1thread-ants-registrationSyN-fp.png}
	\end{subfigure}

	% FreeSurfer recon-all
	\begin{subfigure}[t]{\textwidth}
		\caption{}
		\label{subfig:hotspots-1thread-freesurfer-reconall}
		\includegraphics[width=\textwidth]{figures/hotspots-1thread-freesurfer-reconall.png}
	\end{subfigure}

	\begin{subfigure}[t]{0.49\textwidth}
		\caption{}
		\label{subfig:hotspots-32threads-freesurfer-reconall}
		\includegraphics[width=\textwidth]{figures/hotspots-32threads-freesurfer-reconall.png}
	\end{subfigure}
	\hfill
	% FSL FAST
	\begin{subfigure}[t]{0.49\textwidth}
		\caption{}
		\label{subfig:hotspots-fsl-fast}
		\includegraphics[width=\textwidth]{figures/hotspots-1thread-fsl-fast.png}
	\end{subfigure}

>>>>

\section{Methods}
\subsection{Infrastructure}
For our experiments, we used the \textit{Slashbin cluster} at Concordia University. The cluster has eight compute nodes, each configured with two 16-core Intel(R) Xeon(R) Gold 6130 CPU @ \SI{2.10}{\giga\hertz}, \SI{250}{\gibi\byte} of RAM, \SI{126}{\gibi\byte} of tmpfs, six \SI{447}{\gibi\byte} SSDs with the XFS file system (no RAID enabled), CentOS~8.1 and Linux kernel \textit{4.18.0-477.10.1.el8\_lustre.x86\_64}. \TG{I would defer this subsection to later in the methods as it's not the most central one}

@@

\begin{figure*}[ht!]
	\centering
	% ANTS brainExtraction
	\begin{subfigure}[t]{0.49\textwidth}
		\caption{}
		\label{subfig:hotspots-ants-brainExtraction}
		\includegraphics[width=\textwidth]{figures/hotspots-1thread-ants-brainExtraction.png}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.49\textwidth}
		\caption{}
		\label{subfig:hotspots-ants-brainExtraction-fp}
		\includegraphics[width=\textwidth]{figures/hotspots-1thread-ants-brainExtraction-fp.png}
	\end{subfigure}
	% ANTS registrationSyN
	\begin{subfigure}[t]{0.49\textwidth}
		\caption{}
		\label{subfig:hotspots-ants-registrationSyN}
		\includegraphics[width=\textwidth]{figures/hotspots-1thread-ants-registrationSyN.png}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.49\textwidth}
		\caption{}
		\label{subfig:hotspots-ants-registrationSyN-fp}
		\includegraphics[width=\textwidth]{figures/hotspots-1thread-ants-registrationSyN-fp.png}
	\end{subfigure}
	\caption{ANTS: Single vs. Double precision}
	\label{fig:ants-fp-comparison}
\end{figure*}

\section{old Results}
% TODO
% 2-step approach:
% 1. Which pipeline constitute the long execution time in preprocessing?
% 2. For those, which functions take the most time to execute? (hotspots)
% Any key finding: common point between functions, mem bound, fops, etc.

\subsection{Hotspots}
Figure~\ref{fig:hotspots-fmriprep-anat-subworkflow} depicts the hotspot observed from profiling the fMRIPrep anatomical sub-workflows. In each sub-figure, the x-axis shows the function ID grouped by parent module, which are ordered by CPU time of (1) the modules, then (2) the functions within a module. The function ID are found in the supplementary material Section~\ref{sec:supplementary}. The bars show the average CPU time in seconds of a function and the error bars indicate the standard deviation, with the left y-axis showing the scale. The light-green dots represent the functions cumulative percentage makespan of the application, with the right y-axis showing the scale. The color of the bar indicate the percentage of calls, for the function, that were ``stalled due to demand memory loads and stores''\footnote{\href{https://www.intel.com/content/www/us/en/docs/vtune-profiler/user-guide/2023-0/cpu-metrics-reference.html\#MEMORY-BOUND}{Vtune Memory Bound metric}}, with the left color bar showing the scale.

Similarly, figure~\ref{subfig:hotspots-ants-brainExtraction-fp} shows the same two function accounting for the most time, although in flipped order. Calls from function~2 are memory bound $36.30\pm{4.98}$ percent of the time. Compared to the double precision version of ANTS brainExtraction, the third function runtime is significantly larger. It is an ITK function transforming points while calculating the displacement field of the image.

% TODO FreeSurfer 1-thread sub-figure
% Note:
% - Way more functions
% - Top 5 accounts for only 15% of total time. Much less than other applications.
% - Way more memory bound that previous application.
% - Focusing on Top 5 and memory bound could give good performance boost.

\begin{figure*}[ht!]
	\centering
	% % ANTS brainExtraction
	% \begin{subfigure}[t]{0.49\textwidth}
	% 	\caption{}
	% 	\label{subfig:hotspots-ants-brainExtraction}
	% 	\includegraphics[width=\textwidth]{figures/hotspots-1thread-ants-brainExtraction.png}
	% \end{subfigure}
	% \hfill
	% \begin{subfigure}[t]{0.49\textwidth}
	% 	\caption{}
	% 	\label{subfig:hotspots-ants-brainExtraction-fp}
	% 	\includegraphics[width=\textwidth]{figures/hotspots-1thread-ants-brainExtraction-fp.png}
	% \end{subfigure}

	% % ANTS registrationSyN
	% \begin{subfigure}[t]{0.49\textwidth}
	% 	\caption{}
	% 	\label{subfig:hotspots-ants-registrationSyN}
	% 	\includegraphics[width=\textwidth]{figures/hotspots-1thread-ants-registrationSyN.png}
	% \end{subfigure}
	% \hfill
	% \begin{subfigure}[t]{0.49\textwidth}
	% 	\caption{}
	% 	\label{subfig:hotspots-ants-registrationSyN-fp}
	% 	\includegraphics[width=\textwidth]{figures/hotspots-1thread-ants-registrationSyN-fp.png}
	% \end{subfigure}

	% % FreeSurfer recon-all
	% \begin{subfigure}[t]{\textwidth}
	% 	\caption{}
	% 	\label{subfig:hotspots-1thread-freesurfer-reconall}
	% 	\includegraphics[width=\textwidth]{figures/hotspots-1thread-freesurfer-reconall.png}
	% \end{subfigure}

	% \begin{subfigure}[t]{0.49\textwidth}
	% 	\caption{}
	% 	\label{subfig:hotspots-32threads-freesurfer-reconall}
	% 	\includegraphics[width=\textwidth]{figures/hotspots-32threads-freesurfer-reconall.png}
	% \end{subfigure}
	\hfill
	% FSL FAST
	\begin{subfigure}[t]{0.49\textwidth}
		\caption{}
		\label{subfig:hotspots-fsl-fast}
		\includegraphics[width=\textwidth]{figures/hotspots-1thread-fsl-fast.png}
	\end{subfigure}
